Verification of Stencil Computations through Spatial Specifications

### 2016-07-07 paper.tex
fix use of subfigure
more related work

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{stmaryrd}
\usepackage{color}
\usepackage{graphics}
\usepackage{fancyvrb}
\usepackage{subfigure}
\usepackage{amsthm}
\usepackage{multirow}
\usepackage{minted}
\usepackage{breqn}
%\usepackage{subcaption}
\usepackage{subfigure}
\usepackage{siunitx} % For pretty-printing numeric values and SI units
                     % of measure. e.g., the tabular column type S is
                     % used to print nice-looking tables of numbers.
\sisetup{ % defaults
  group-separator={,},
  group-minimum-digits={3},
  output-decimal-marker={.},
  table-format = 6
}

@@

\begin{figure}[t]
  \centering
  %\subcaptionbox{\label{fig:jacobicorrect}}{\includegraphics[width=0.1\textwidth]{jacobi-correct.png}}
  %\subcaptionbox{}{\includegraphics[width=0.1\textwidth]{jacobi-incorrect1.png}}
  %\subcaptionbox{}{\includegraphics[width=0.1\textwidth]{jacobi-incorrect2.png}}
  %\subcaptionbox{}{\includegraphics[width=0.1\textwidth]{jacobi-incorrect3.png}}
  \subfigure[]{\label{fig:jacobicorrect}\includegraphics[width=0.1\textwidth]{jacobi-correct.png}}
  \subfigure[]{\includegraphics[width=0.1\textwidth]{jacobi-incorrect1.png}}
  \subfigure[]{\includegraphics[width=0.1\textwidth]{jacobi-incorrect2.png}}
  \subfigure[]{\includegraphics[width=0.1\textwidth]{jacobi-incorrect3.png}}
  \caption{Greyscale output of several perturbations of a Jacobi
    iteration program. The correct output is shown in
    (\subref{fig:jacobicorrect}).}
  \label{fig:jacobi}
\end{figure}

@@

% Abe, 2013, Model checking stencil computations
% -- Unable to obtain copy
% In an alternate approach,~\citet{Abe:2013:IPDPSW}

@@

% GPGPU
This work shares some similarities with efforts to verify kernels
written for General-Purpose Graphics Processing Unit (GPGPU)
programming, such as in \citet{Blom:2014:SoCP}. Stencils are a form of
kernel, and GPGPU programming can be viewed as a massively parallel
method of transforming a large matrix. However, that work focuses
mainly on the synchronisation of such kernels, and the avoidance of
data races, while we are interested in the correctness of stencils
embedded within a more typical general-purpose programming language.

@@

% Pochoir
\citet{Tang:2011:SPAA} defines a specification language for writing
stencils embedded in C++ (with Cilk extensions) that are then compiled
into fast parallel programs based on trapezoidal decompositions with
hyperspace cuts. The Pochoir specification language is used for
describing the computing kernel, boundary conditions and shape of the
stencil. Programs are first compiled in C++ with a template library
that checks whether the annotations are used correctly, and then a
Pochoir preprocessor may be applied to generate the high-performance
parallel version that takes advantage of the Cilk compiler.  Pochoir
is aimed primarily at assisting programmers who may be reluctant to
implement the tricky cache-oblivious ``hypertrapezoidal'' algorithms
that are significantly faster than the typical looping implementations
of stencil computations.

% 3D stencil codes
%\citet{Zhang:2012:CGO}

>>>>

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{stmaryrd}
\usepackage{color}
\usepackage{graphics}
\usepackage{fancyvrb}
\usepackage{subfigure}
\usepackage{amsthm}
\usepackage{multirow}
\usepackage{minted}
\usepackage{breqn}
%\usepackage{subcaption}
\usepackage{siunitx} % For pretty-printing numeric values and SI units
                     % of measure. e.g., the tabular column type S is
                     % used to print nice-looking tables of numbers.
\sisetup{ % defaults
  group-separator={,},
  group-minimum-digits={3},
  output-decimal-marker={.},
  table-format = 6
}

@@

\begin{figure}[t]
  \centering
  %\subcaptionbox{\label{fig:jacobicorrect}}{\includegraphics[width=0.1\textwidth]{jacobi-correct.png}}
  %\subcaptionbox{}{\includegraphics[width=0.1\textwidth]{jacobi-incorrect1.png}}
  %\subcaptionbox{}{\includegraphics[width=0.1\textwidth]{jacobi-incorrect2.png}}
  %\subcaptionbox{}{\includegraphics[width=0.1\textwidth]{jacobi-incorrect3.png}}
  \subfigure[]{\label{fig:jacobicorrect}\includegraphics[width=0.1\textwidth]{jacobi-correct.png}}
  \subfigure[]{\includegraphics[width=0.1\textwidth]{jacobi-incorrect1.png}}
  \subfigure[]{\includegraphics[width=0.1\textwidth]{jacobi-incorrect2.png}}
  \subfigure[]{\includegraphics[width=0.1\textwidth]{jacobi-incorrect3.png}}
  \caption{Greyscale output of several perturbations of a Jacobi
    iteration program. The correct output is shown in
    \subref{fig:jacobicorrect}.}
  \label{fig:jacobi}
\end{figure}

@@

@@

% GPGPU
This work shares some similarities with efforts to verify kernels
written for General-Purpose Graphics Processing Unit (GPGPU)
programming, such as in \citet{Blom:2014:SoCP}. Stencils are a form of
kernel, and GPGPU programming can be viewed as a massively parallel
method of transforming a large matrix. However, that work focuses
mainly on the synchronisation of such kernels, and the avoidance of
data races, while we are interested in the correctness of stencils
embedded within a more typical general-purpose programming
language. Other work on GPGPU computation, such as
\citet{Zhang:2012:CGO}, has been primarily focused on generating
optimised code based upon on relatively simple specifications: to
provide performance while keeping the programmer's effort within
reason.

@@

% Pochoir
\citet{Tang:2011:SPAA} defines a specification language for writing
stencils embedded in C++ (with Cilk extensions) that are then compiled
into fast parallel programs based on trapezoidal decompositions with
hyperspace cuts. The Pochoir specification language is used for
describing the computing kernel, boundary conditions and shape of the
stencil. Programs are first compiled in C++ with a template library
that checks whether the annotations are used correctly, and then a
Pochoir preprocessor may be applied to generate the high-performance
parallel version that takes advantage of the Cilk compiler.  Pochoir
is aimed primarily at assisting programmers who may be reluctant to
implement the tricky cache-oblivious ``hypertrapezoidal'' algorithms
that are significantly faster than the typical looping implementations
of stencil computations. Like much of the related work, the main goal
is optimisation rather than catching bugs and verification of
program correctness.

By contrast, the work of \citet{Abe:2013:IPDPSW} has looked at the
correctness problem explicitly by bringing a form of model-checking to
verify certain stencil computations in partitioned global address
space languages. These are scenarios where an array is divided into
subarrays on multiple processors but each has the ability to access
the others' memory. The authors avoid the state explosion dilemma of
model-checking by relying upon the fact that accesses to the
``boundary elements'' will require a different method by virtue of the
non-locality of that memory. Since this access is performed in a
different manner, those can be identified automatically by static
analysis. To achieve this, the authors have designed a new language to
be used for writing stencil computations. Much of the specification
effort goes towards describing the nuts and bolts of distributing the
computation over multiple processors. The code for the stencil kernel
is generated from a relatively high-level specification. This approach
differs from Camfort because we are interested in integrating directly
into existing, legacy codebases and long-established languages such as
older versions of Fortran. We infer stencil specifications from
Fortran code and check annotations on stencil computations within
Fortran code.
